Phylodynamic inference of contact network parameters through approximate Bayesian computation

### 2016-04-16 doc/thesis/appendix.tex
More writing yay

>>>>

\begin{figure}
  \centering
  \includegraphics{gridsearch-alpha-kernel.pdf}
  \caption{                                              
    Grid search kernel scores for testing trees simulated under various             
    \gls{alpha} values. The other \gls{BA} parameters were fixed at \gls{I} =
    1000, \gls{N} = 5000, and \gls{m} = 2. 
  }        
  \label{fig:gridalpha}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{gridsearch-I-kernel.pdf}
  \caption{                                              
    Grid search kernel scores for testing trees simulated under various             
    \gls{I} values. The other \gls{BA} parameters were fixed at \gls{alpha} =
    1.0, \gls{N} = 5000, and \gls{m} = 2. 
  }        
  \label{fig:gridI}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{gridsearch-m-kernel.pdf}
  \caption{                                              
    Grid search kernel scores for testing trees simulated under various             
    \gls{m} values. The other \gls{BA} parameters were fixed at \gls{alpha} =
    1.0, \gls{I} = 1000, and \gls{N} = 5000.
  }        
  \label{fig:gridm}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{gridsearch-N-kernel.pdf}
  \caption{                                              
    Grid search kernel scores for testing trees simulated under various             
    \gls{N} values. The other \gls{BA} parameters were fixed at \gls{alpha} =
    1.0, \gls{I} = 1000, and \gls{m} = 2. 
  }        
  \label{fig:gridN}
\end{figure}

### 2016-04-16 doc/thesis/conclusion.tex
More writing yay

Our work had three main aims. First, we developed a method to estimate contact
network model parameters from phylogenetic data. This method widens the field
of models which may be investigated in a phylodynamic context. Second, we
investigated the parameters of the \gls{BA} model. We determined through
simulation studies that the preferential attachment power $\alpha$ and number
of infected nodes $I$ had a substantial impact on transmission tree shape, and
could be estimated using our method.

>>>>

Our work had three main aims. First, we developed a method to estimate contact
network model parameters from phylogenetic data, based on kernel approximate
Bayesian computation. This method widens the field of models which may be
investigated in a phylodynamic context.

Our second aim was to investigate the parameters of the \gls{BA} model, in
terms of their impact on tree shape and our ability to reconstruct them with
kernel-\gls{ABC}. We determined through simulation studies that the
preferential attachment power $\alpha$ and number of infected nodes $I$ had a
substantial impact on transmission tree shape, and could be estimated using our
method.

The final aim of our work was to apply our method to real world \gls{HIV}
datasets. This analysis indicated heterogeneity in the contact network
structures underlying several distinct local epidemics. The five datasets
analysed fell into three categories (Figure~\ref{fig:abchpd}). First, we
estimated a preferential attachment power between 0.5 and 1 for the epidemics
studied by \textcite{cuevas2009hiv} and \textcite{li2015hiv}, with credible
intervals occupying nearly the entire region from 0 to 1.
\citeauthor{cuevas2009hiv} studied a group of newly diagnosed individuals in
the Basque Country, Spain. Although the individuals were of mixed risk groups,
and therefore unlikely to comprise a single contact network, a high proportion
of them (47\%) grouped into local clusters based on genetic distance. The low
estimated attachment power for these data is consistent with the sampled
sequences comprising many distinct sub-networks rather than a single connected
network. \citeauthor{li2015hiv} sampled a large number of acutely infected MSM
in Shanghai, China, in which we identified a large cluster from the phylogeny
using a patristic distance cutoff~\autocite{poon2015impact}. The low attachment
power estimated for this dataset was surprising given the high phylogenetic
relatedness of the sequences. It is possible that the number and diversity of
circulating recombinant forms in the data introduced errors into our estimated
viral phylogeny.

For the outbreaks studied by \textcite{niculescu2015recent} and
\textcite{wang2015targeting}, the estimated $\alpha$ was close to one, with a
narrower credible interval than for the other studies.
\citeauthor{niculescu2015recent} studied a recent outbreak among Romanian
injection drug users (IDU), while \citeauthor{wang2015targeting} sampled
acutely infected MSM in Beijing, China. Both studies discovered a high degree
of phylogenetic relatedness owing to the recent infection times and homogeneous
risk groups of the studied populations. The estimated number of infections for
these datasets were also quite low, although the HPD interval for
\citeauthor{wang2015targeting} was much wider than that for
\citeauthor{niculescu2015recent}.

The final studied dataset was an outlier in terms of estimated parameters.
\textcite{novitsky2013phylogenetic} sampled approximately 44\% of the
HIV-infected individuals in the northern area of Mochudi, Botswana. Additional
sampling in a later study~\autocite{novitsky2014impact} brought the genotyping
coverage up to 70\%. Even with such a high sampling coverage, we did not detect
any large clusters using patristic distance, and therefore chose to analyze a
subtree instead. Estimates of $\alpha$ and $N$ both had very wide HPD
intervals and were markedly different from the other datasets. The estimated
number of infected nodes was also extremely high, much higher than the
estimated HIV prevalence of the town. Several factors may have contributed to
these results. First, the authors note that the their sample was 75\% female.
In a primarily heterosexual risk environment, removal of a disproportionate
number of males from the network could obfuscate the true network structure,
for example if the majority of highly connected nodes were of one gender.
Second, the town in question was in close proximity to the country's capital,
and the authors indicated that a high amount of migration takes place between
the two locations. This suggests that the contact network may include a much
larger group based in the capital city, which would explain the high estimate
of $I$.

### 2016-04-16 doc/thesis/intro.tex
More writing yay

One mechanism which has been shown to lead to scale-free networks is
\defn{preferential attachment}~\autocite{simon1955class, barabasi1999emergence}.
Under this process, networks are formed by starting with a small number $m_0$
of nodes. New nodes are added one at a time until there are a total of $N$ in
the network. Each time a new node is added, $m \geq 1$ edges are added from it
to other nodes in the graph. In the original
formulation~\autocite{barabasi1999emergence}, the partners of the new node are
chosen with probability linearly proportional to their degree. However,
\citeauthor{barabasi1999emergence} suggest extending the model such that the
probability of choosing a partner of degree $d$ is proportional to $d^\alpha$
for some constant $\alpha$, and we use this extension here.

>>>>

One mechanism which has been shown to lead to scale-free networks is
\defn{preferential attachment}~\autocite{simon1955class,
barabasi1999emergence}. The simplest preferential attachment model is known as
the \acrlong{BA} (BA) model after its
inventors~\autocite{barabasi1999emergence}. Under this model, networks are
formed by starting with a small number $m_0$ of nodes. New nodes are added one
at a time until there are a total of $N$ in the network. Each time a new node
is added, $m \geq 1$ edges are added from it to other nodes in the graph. In
the original formulation, the partners of the new node are chosen with
probability linearly proportional to their degree. However,
\citeauthor{barabasi1999emergence} suggest extending the model such that the
probability of choosing a partner of degree $d$ is proportional to $d^\alpha$
for some constant $\alpha$, and we use this extension here. When $m = 1$, the
network takes on the distinctive shape of a tree, that is, it does not contain
any cycles. Cycles are present in the network for all all other $m$ values.

### 2016-04-16 doc/thesis/methods.tex
More writing yay

As in the previous section, we will begin by describing a single experiment,
and then list the variables for which similar experiments were performed. We
varied \gls{alpha} along a narrowly spaced grid of values: 0, 0.01, \ldots, 2.
For each value, fifteen networks were generated with \software{igraph}, and
transmission trees were simulated over each using \software{nettree}. These
trees will be referred to as ``grid trees'', and their associated values ``grid
values''. Next, one further test tree was simulated with the test value
\gls{alpha} = 0. Both the grid trees and the test tree had 500 tips, and were
simulated with the other \gls{BA} parameters set to $N$ = 5000, $m$ = 2, and
$I$ = 1000. The test tree was compared to each of the grid trees using the tree
kernel, with the meta-parameters set to $\lambda = 0.3$ and $\sigma = 4$, using
the \software{treekernel} binary. The median kernel score was calculated for
each grid value. Grid values were resampled with probability proportional to
their median kernel scores. A point estimate for the test value was obtained by
taking the highest point of an estimated kernel density of the resampled grid
values, and a 95\% credible interval was obtained using the
\software{HPDinterval} function in the \software{coda} package.

>>>>

As in the previous section, we will begin by describing a single experiment,
and then list the variables for which similar experiments were performed. We
varied \gls{alpha} along a narrowly spaced grid of values: 0, 0.01, \ldots, 2.
For each value, fifteen networks were generated with \software{igraph}, and
transmission trees were simulated over each using \software{nettree}. These
trees will be referred to as ``grid trees'', and their associated values ``grid
values''. Next, one further test tree was simulated with the test value
\gls{alpha} = 0. Both the grid trees and the test tree had 500 tips, and were
simulated with the other \gls{BA} parameters set to $N$ = 5000, $m$ = 2, and
$I$ = 1000. The test tree was compared to each of the grid trees using the tree
kernel, with the meta-parameters set to $\lambda = 0.3$ and $\sigma = 4$, using
the \software{treekernel} binary. The median kernel score was calculated for
each grid value, and the scores were normalized such that the area under the
curve was equal to 1. The grid value with the highest median kernel score was
taken as the point estimate for the test value, and a 95\% credible interval
was obtained using the \software{hpd} function in the \software{TeachingDemos}
package.

### 2016-04-16 doc/thesis/results.tex
More writing yay

\subsection{Separable parameters}

Our first simulation-based experiment was designed to determine which network
model parameters were potentially estimable by kernel-ABC, and to find the best
set of kernel meta-parameters to carry through to future experiments. For each
parameter of the \gls{BA} model, we simulated trees under three distinct
values, and computed the cross-validation accuracy of a kernel-\gls{SVM}
classifier for the parameter.

In Figure~\ref{fig:patrees}, we show that trees simulated under different
values of \gls{alpha} are visibly quite distinctive. In particular, higher
values of \gls{alpha} result in networks with a small number of highly
connected nodes (see Figure~\ref{fig:pabounds}) which, once infected, are
likely to transmit to many other nodes. This results in a more unbalanced,
ladder-like structure in the phylogeny. Kernel-\gls{PCA} projections show that
all three \gls{alpha} values are well separated from each other under several
different sampling scenarios (Figure~\ref{fig:pakpca}). With smaller trees,
it becomes harder to visually disinguish \gls{alpha} = 0.5 from \gls{alpha} =
1.0 using the first two principal components.

With suitably chosen meta-parameters, the accuracy of the kernel-\gls{SVM}
classifier for \gls{alpha} was very high under a variety of prevalence and
sampling scenarios (Figure~\ref{fig:pacrossv}). Accuracy was highest, with
$R^2$ values above 0.95, for the largest trees and complete sampling (bottom
center panel). However, even for trees of size 100 sampled from an epidemic on
2000 nodes, the $R^2$ was above 0.8 (top right panel). In all cases, the
accuracy of a Sackin's index-based classifier was also quite high, at about
0.75. An \gls{SVM} classifier using only the \gls{nltt}
statistic~\autocite{janzen2015approximate} was slightly worse than Sackin's
index, with an $R^2$ of X.

The fact that Sackin's index is informative of \gls{alpha} was
unsurprising, given the clear differences in tree balance observed under
different \gls{alpha} values.

\begin{figure}[ht]
  \centering
  \label{fig:patrees}
  \includegraphics{kernel_pa_trees}
  \caption[Visibly distinctive trees simulated under three values of \gls{alpha}]{
    Epidemics of 1000 infected were simulated on \gls{BA} networks of 5000
    nodes, with \gls{alpha} equal to 0.5, 1.0, or 1.5. Transmission trees were
    created by sampling 500 infected nodes. Higher \gls{alpha} values produce
    networks with a small number of highly-connected nodes, which results in a
    highly unbalanced, ladder-like tree structure.
  }
\end{figure}

\begin{figure}[ht]
  \centering
  \label{fig:pacrossv}
  \includegraphics{kernel_pa_crossv}
  \caption[Cross-validation performance of kernel support vector machine
  classifier for preferential attachment power]{
    Cross-validation performance of kernel support vector machine classifier
    for preferential attachment power.
  }
\end{figure}

\begin{figure}[ht]
  \centering
  \label{fig:pakpca}
  \includegraphics{kernel_pa_kpca}
  \caption[Projection of kernel matrix for different attachment power values
  onto its first two principal components]{
    Projection of the kernel matrix for different preferential attachment power
    values onto its first two principal components, for eight simulation
    scenarios. Each point corresponds to a simulated transmission tree, and is
    coloured by preferential attachment power. Facets are number of infected
    nodes (horizontal), and number of sampled tips (vertical). The parameters
    to the tree kernel were $\lambda = 0.3$ and $\sigma = 4$, and the nLTT was
    not used. Qualitatively, trees with a larger number of tips are easier to
    separate in kernel space, regardless of what sampling proportion they
    represent. In all cases, the highest attachment power can be separated from
    the other two, but the two lowest values become hand to distinguish with in
    the 100-tip trees.
  }
\end{figure}

Figure~\ref{fig:rsquared} shows the cross-validation accuracy of classifiers
for all four parameters of the \gls{BA} model. The classifier for \gls{I} also
exhibited high performance, although not quite as high as that for \gls{alpha}.
The average $R^2$ values were 0.70 for 100-tip trees, or 0.93 for 500-tip
trees. On the other hand, $m$ and $N$ were much harder to classsify. For $m$,
all $R^2$ values were below 0.5, falling to virtually zero in four scenarios.
The $R^2$ values for $N$ were highly variable, ranging from 0.08 to 0.82
depending on $I$ and the number of tips in the tree. In almost all cases, both
the nLTT and Sackin's index classifiers were worse than the tree kernel, except
for classifying $N$ with $I = 2000$ and $100$ tips, where the nLTT gave a small
advantage (see the supplements). The greatest disadvantage for the Sackin's
index-based classifier was on the $I$ parameter, where it had an $R^2$ of less
than 0.1, despite the relatively high accuracy of the tree kernel (see the
supplements).

\begin{figure}[ht]
  \centering
  \includegraphics{kernel_rsquared}
  \caption[Cross-validation accuracy of kSVM classifiers for \gls{BA}
  model parameters.]{Cross-validation $R^2$ values of kernel-SVM classifier for
    \gls{BA} model parameters. Points and ranges are median and 95\% quantiles
    over 1000 2-fold cross-validations.}
  \label{fig:rsquared}
\end{figure}

% needs to go in methods

%It is important to note that, if we cut off the epidemic simulation when 500
%nodes are infected, the resulting tree will be shorter (in calendar time) than
%if we continue until 2000 nodes are infected. However, this information is not
%used when building the classifier, since the branch lengths in each tree are
%scaled by their mean. Therefore, the high performance of the classifier is due
%to structural differences captured by the tree kernel, rather than the trees
%simply having different heights.

\subsection{Accuracy of marginal estimates}

We used grid search to obtain \defn{marginal} estimates for each network
parameter while holding all other parameters fixed. We observed that kernel
scores were highest at the values of \gls{alpha} on the grid closest to the
true values, as shown in Figure~\ref{fig:gridkernel}. However, there was a much
stronger spike in kernel scores near the true value for \gls{alpha} = 1.0 and
1.25. This is recapitulated when we look at the accuracy of point estimates
obtained by taking the grid value with the highest median kernel score. As
shown in Figure~\ref{fig:gridest}, while the estimates are generally close to
the true value, they are much closer for \gls{alpha} = 1.25 than for the other
values.

\begin{figure}[ht]
  \centering
  \includegraphics{gridsearch_pa_kernel}
  \caption[Grid search kernel scores]{
    Grid search kernel scores for testing trees simulated under various
    \gls{alpha} values. All epidemics had \gls{I} = 1000 infected nodes, on
    \gls{BA} networks of size \gls{N} = 5000 with \gls{m} fixed at 2. Colours
    indicate the number of sampled tips.
  }
  \label{fig:gridkernel}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics{gridsearch_pa_estimates}
  \caption[Marginal estimates of \gls{alpha} obtained with grid search]{
    Marginal estimates of \gls{alpha} obtained with grid search. Training trees
    were simulated on a narrowly spaced grid of \gls{alpha} values, and
    compared to testing trees using the tree kernel. The \gls{alpha} value in
    the grid with the highest median kernel score was taken as the point
    estimate for the testing tree. These point estimates are shown as black
    dots. The dashed line is the identity.
  }
  \label{fig:gridest}
\end{figure}

@@

>>>>

\subsection{Kernel classifiers}

Trees simulated under different values of \gls{alpha} were visibly quite
distinct (\cref{fig:alphatrees}). In particular, higher values of \gls{alpha}
produce networks with a small number of highly connected nodes which, once
infected, are likely to transmit to many other nodes. This results in a more
unbalanced, ladder-like structure in the phylogeny, compared to networks with
lower \gls{alpha} values. None of the other three parameters produced trees
which were as easily distinguished from each other
(\cref{fig:Itrees,fig:mtrees,fig:Ntrees,fig:Itrees}).
Sackin's index, which measures tree imbalance, was significantly correlated with
all four parameters
    (for $\alpha$, $I$, $m$, and $N$ respectively: Spearman's rho =
     0.85,
     \ensuremath{-0.12},
     \ensuremath{-0.13},
     0.09;
     $p$-values
     $<10^{-5}$,
     $0.003$,
     $<10^{-5}$,
     $<10^{-5}$)
The ratio of internal to terminal branch lengths was negatively correlated with
\gls{alpha} and \gls{I}, and positively corelated with \gls{N} and \gls{m}
  (Spearman's rho
    \ensuremath{-0.84},
    \ensuremath{-0.69},
    0.1,
    0.18;
  $p$-values
    $0$,
    $2\!\times\!10^{-85}$,
    $4\!\times\!10^{-7}$,
    $0$).

\begin{figure}[ht]
  \centering
  \includegraphics{kernel-alpha-tree.pdf}
  \caption[Visibly distinctive trees simulated under three values of \gls{alpha}]{
    Epidemics simulated on \gls{BA} networks of 5000 nodes, with \gls{alpha}
    equal to 0.5, 1.0, or 1.5, until 1000 individuals were infected.
    Transmission trees were created by sampling 500 infected nodes. Higher
    \gls{alpha} values produced networks with a small number of
    highly-connected nodes, resulting in highly unbalanced, ladder-like trees.
  }
  \label{fig:alphatrees}
\end{figure}

\Cref{fig:kpca} shows \gls{kPCA} projections of the simulated trees onto the
first two principal components of the kernel matrix. The figure shows only the
simulations with 500-tip trees and, for all parameters except $I$, 1000
infected nodes. The three \gls{alpha} and \gls{I} values considered are well
separated from each other in feature space. On the other hand, the three
\gls{N} values overlap significantly, and the three \gls{m} values are
virtually indistinguishable. Similar observations can be made for other values
of \gls{I} and the number of tips
(\cref{fig:alphakpca,fig:Nkpca,fig:Ikpca,fig:mkpca}). The values of \gls{I} and
\gls{N} separate more clearly with larger numbers of tips, and in the case of
\gls{N} larger epidemic sizes.

\begin{figure}[ht]
  \centering
  \includegraphics{kernel-kpca.pdf}
  \caption[\gls{kPCA} projections of simulated trees under varying \gls{BA}
           parameter values]{
    Each parameter of the \gls{BA} model was individually varied to produce 300
    simulated trees. Kernel matrices were formed from all pairwise kernel
    scores among each set of 300 trees. The trees were projected onto the first
    two principal components of the kernel matrix calculated using \gls{kPCA}.
    All trees had 500 tips. The parameters not being varied were set to
    \gls{alpha} = 1, \gls{I} = 1000, \gls{m} = 2, and \gls{N} = 5000. The tree
    kernel meta-parameters were $\lambda = 0.3$ and $\sigma = 4$.
  }
  \label{fig:kpca}
\end{figure}

Accuracy of the \gls{kSVM} classifiers varied based on the parameter being
tested (\cref{fig:rsquared}, left). Classifiers based on two other tree
statistics, the \gls{nltt} and Sackin's index, generally exhibited worse
performance than the tree kernel, although the magnitude of the disparity
varied between the parameters (\cref{fig:rsquared}, centre and right). The
results were largely robust to variations in the tree kernel meta-parameters
$\lambda$ and $\sigma$, although accuracy varied between different epidemic and
sampling scenarios
(\cref{fig:alphacrossv,fig:mcrossv,fig:Icrossv,fig:Ncrossv}).

When classifying $\alpha$, the kernel-SVM classifier had an average $R^2$ of 
    0.92,
compared to 
    0.56
for the \gls{nltt}-based SVM, and
    0.75
for the linear regression against Sackin's index. There was little variation
about the mean for different tree and epidemic sizes. No classifier could
accurately identify the $m$ parameter in any epidemic scenario, with average
$R^2$ values of 
  0.12 for \gls{kSVM},
  0.01 for the \gls{nltt}, and
  0.06
for Sackin's index. Again, there was little variation in accuracy between
epidemic scenarios, although the accuracy of the \gls{kSVM} was slightly higher
on 1000-tip trees.

The accuracy of classifiers $I$ varied significantly with the number of tips in
the tree. For 100-tip trees, the average $R^2$ values were
  0.7,
  0.55, and
  0.02
for the tree kernel, \gls{nltt}, and Sackin's index respectively. For 500-tip
trees, the values increased to
  0.93,
  0.83, and
  0.07.
Finally, the performance of classifiers for $N$ depended heavily on the
epidemic scenario. The $R^2$ of the \gls{kSVM} classifier ranged from
  0.08
for the smallest epidemic and smallest sample size, to
  0.82
for the largest. Likewise, $R^2$ for the \gls{nltt}-based SVM ranged from 
  0.01
to
  0.54.
Sackin's index did not accurately classify $N$ in any scenario, with an average
$R^2$ of
  0.03
and little variation between scenarios.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{kernel-rsquared.pdf}
  \caption{
      Cross-validation accuracy of kernel-SVM classifier (left), SVM classifier
      using \gls{nltt} (centre), and linear regression using Sackin's index
      (right) for \gls{BA} model parameters. Kernel meta-parameters were set to
      $\lambda = 0.3$ and $\sigma = 4$. Each point was calculated based on 300
      simulated transmission trees over networks with three different values of
      the parameter being tested. Vertical lines are empirical 95\% confidence
      intervals based on 1000 two-fold cross-validations.
  }
  \label{fig:rsquared}
\end{figure}

\subsection{Grid search}

The accuracy of grid search estimates largely paralleled that of the \gls{kSVM}
classifiers. \Cref{fig:gridest} shows point estimates and 95\% highest density
intervals for each of the \gls{BA} parameters, for one replicate experiment
with 500-tip trees. For all parameters except $m$, the error of point estimates
was negatively correlated with the number of sampled tips in the tree (for
\gls{alpha}, \gls{I}, and \gls{N} respectively: Spearman's $\rho$ = 
    \ensuremath{-0.22},
    \ensuremath{-0.41},
    \ensuremath{-0.16};
$p$-value = 
    $4\!\times\!10^{-4}$,
    $2\!\times\!10^{-4}$,
    $0.01)$.
The highest density intervals obtained for all parameters were extremely wide,
occcupying $>$90\% of the grid in all cases.

The \gls{alpha} parameter was the most accurately estimated, with point
estimates having an average deviation of 
    0.14
from the true value, on a grid from 0 to 2. The error was negatively correlated
with the true value of \gls{alpha} 
    (Spearman's $\rho$ = \ensuremath{-0.26},
     $p = 1\!\times\!10^{-5}$),
although the relationship was clearly nonlinear (\cref{fig:gridest,fig:gridalpha}).
The accuracy was highest for the test value \gls{alpha} = 1.25
    (mean error 0.02)
which exhibited markedly different behaviour than the other values in terms of
the distribution of kernel scores along the grid (\cref{fig:gridalpha}). In
particular, there was a very pronounced peak in scores around the true value,
in contrast to most other values where the scores were flat around the true
value. The peak was also observed to a lesser degree for \gls{alpha} = 1. The
average absolute error of the point estimates for \gls{I} was 
    298,
on a grid of 500 to 5000, and the errors were not significantly correlated with
the true value of \gls{I}. Kernel score distributions for all test values
exhibited the same rounded shape (\cref{fig:gridI}).

For the other two parameters, \gls{m} and \gls{N}, accuracy was much poorer
(\cref{fig:gridest,fig:gridm,fig:gridN}). The average error for \gls{m} was
    1.31,
on a grid from 1 to 6; this error was positively correlated with increasing
\gls{m}
    (Spearman's $\rho$ = 0.25,
     $p = 6\!\times\!10^{-4}$).
This positive correlation was apparently due to the much lower error for $m =
1$ than for the other $m$ values (\cref{fig:gridm})
    (mean errors 0.1 for
     $m = 1$ vs. 1.5466667 for
     $m > 1$).
The value $m = 1$ causes the network to take on a distinct shape relative to
higher $m$ values, namely a tree (\textit{i.e.} there are no cycles,
\cref{subsec:treeshape}). The average error for \gls{N} was 
    2419,
on a grid from 1000 to 15000, and was positively correlated with the true value
of \gls{N}
    (Spearman's $\rho$ = 0.43,
     $p = 2\!\times\!10^{-12}$).

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{gridsearch-example}
  \caption[Grid search estimates of \gls{BA} model parameters]{Point estimates
      and 95\% highest density intervals for each \gls{BA} model parameter,
      obtained using grid search. Networks and transmission trees were
      simulated over a grid of values for each parameter while holding the
      others fixed. For a subset of the grid values ($x$-axis), test networks
      and trees were created and compared to each tree on the grid using the
      tree kernel. The kernel scores along the grid were normalized to resemble
      a probability distribution, from which the mode and highest density
      interval were calculated. Shown values correspond to one replicate
      experiment, with trees of size 500.
  } 
  \label{fig:gridest}
\end{figure}

@@

We used kernel-\gls{ABC} to estimate the parameters of the \gls{BA} model 
on simulated trees where the true parameter values were known. Point estimates
for each parameter are shown in Figure~\ref{fig:abcpt}. Of the four parameters,
\gls{alpha} was the most accurately estimated, with a median [IQR] absolute error
of 
    0.11 
    [0.05-0.18].
The accuracy of the estimates was not significantly different between values of
$m$ or $I$ (both one-way ANOVA,
    $p$ = 0.1
and 
    0.25),
although the errors when the true value of \gls{alpha} was zero were
significantly greater than the other values 
    (Wilcoxon rank-sum test, $p$ = \ensuremath{6.41\times 10^{-4}}).
The error in the estimated value of $I$ was
    306 
    [108-607].
Errors were significantly higher for $\alpha \geq 1$
    (Wilcoxon rank-sum test, $p$ = \ensuremath{6.12\times 10^{-4}})
and for $I$ = 2000
    ($p$ = \ensuremath{1.58\times 10^{-6}}),
but not for any values of $m$
    (one-way ANOVA, $p$ = 0.33).
The $m$ parameter was estimated correctly in
    37 \%
of simulations, with an error of one in
    40 \%
and of two or more in 
    22 \%
(the only possible $m$ values were 2, 3, 4, or 5). The true values of
$m$ and $I$ did not significantly affect the error
    (one-way ANOVA, $p$ = 0.5 and
                          0.68),
but the accuracy was significantly lower for integral than non-integral
values of \gls{alpha}
    (Wilcoxon rank-sum test, $p$ = \ensuremath{7.2\times 10^{-3}}).
Finally, the total number of nodes \gls{N} was consistently over-estimated by
about a factor of two
    (error \ensuremath{6.59\times 10^{3}} 
    [\ensuremath{4.21\times 10^{3}}-\ensuremath{8.28\times 10^{3}}]).
No other parameters influenced the accuracy of the $N$ estimates 
    (one-way ANOVA, $p \geq$ NA).

\Cref{fig:abcex} shows the \gls{ABC} approximation to the posterior
distribution on the \gls{BA} parameters for one simulation (equivalent plots
for all the simulations can be found in the supplemental materials). \Gls{HPD}
intervals around \gls{alpha} and \gls{I} were narrow relative to the region of
nonzero prior density, whereas the intervals for $m$ and \gls{N} were widely
dispersed. \Cref{tab:abchpd} shows point estimates and 95\% \gls{HPD}
intervals averaged over all simulations.

\begin{table}
    \centering
    \input{\tablepath/abc-hpd}
    \caption{Average widths of 95\% confidence intervals for \gls{BA} model
    parameters estimated with kernel-\gls{ABC}.}
    \label{tab:glm}
\end{table}

\subsection{Characterization of power-law exponent in Barab\'asi-Albert networks}

\Cref{tab:glm} shows the estimated parameters for a log-link \gls{GLM} fitted
to the observed distribution of \gls{gamma} values. The coefficients are
interpretable as multiplicative effects.

\begin{table}
    \centering
    \input{\tablepath/pa-gamma-glm}
    \caption{Estimated \gls{GLM} parameters for relationship between power-law
    exponent \gls{gamma} and \gls{BA} model parameters.}
    \label{tab:glm}
\end{table}

\subsection{Real data}

We applied kernel-ABC to five published HIV datasets (\cref{tab:data}),
and found substantial heterogeneity among the parameter estimates
(\cref{fig:abchpd}). Two of the datasets~\autocite{niculescu2015recent,
wang2015targeting} had estimated $\alpha$ values near unity (MAP estimates
[95\% HPD] 
  1.06 
  [0.63 - 
   1.27]
and
  1 
  [0.41 -
   1.16] respectively). 
Another two datasets~\autocite{li2015hiv, cuevas2009hiv} had lower estimated
values and wider HPD intervals
  (0.77 
  [0.01 - 
  1.03]
and
  0.66 
  [0.03 -
   0.84]). 
The \textcite{novitsky2014impact} data had an extremely low estimated $\alpha$
and a very wide HPD interval
  (0.17 
  [0.04 -
   1.39]). 
For all the datasets except \citeauthor{novitsky2014impact}, estimated values of
$I$ were below 2000, with narrow HPD intervals around two of the
datasets
  (\citeauthor{cuevas2009hiv}, 880 
  [290 -
   \ensuremath{1.51\times 10^{3}}];
   \citeauthor{niculescu2015recent}, 175
  [138 - 
   454])
and wider intervals around the other two
  (\citeauthor{li2015hiv}, \ensuremath{1.59\times 10^{3}} 
  [284 -
   \ensuremath{3.81\times 10^{3}}];
   \citeauthor{wang2015targeting}, 651
  [268 - 
   \ensuremath{4.24\times 10^{3}}]).
The \citeauthor{novitsky2014impact} data was again the outlier, with a very high
estimated $I$, and HPD interval spanning almost the entire prior region
  (\ensuremath{7.55\times 10^{3}} 
  [228 -
   \ensuremath{8.92\times 10^{3}}]).
No information was gleaned about the $m$ parameter, with the HPD interval
occupying the entire prior region for all datasets. The estimates of $N$ were
similarly uninformative, with the exception that the point estimate for the
\citeauthor{wang2015targeting} data was smaller
  (\ensuremath{5.84\times 10^{3}})
than the estimates for other datasets
 (average \ensuremath{8.93\times 10^{3}}).

\begin{figure*}[ht]
  \centering
  \includegraphics{realdata-hpd-bc.pdf}
  \vspace{8pt}
  \caption{
      Maximum \textit{a posteriori} point estimates and 95\% HPD intervals for
      parameters of the BA network model, fitted to five published HIV datasets
      with kernel-ABC.
  }
  \label{fig:abchpd}
\end{figure*}
